Light Curves \& Planet Candidates in Open Clusters from TESS Sectors 6 \& 7

### 2019-07-05 paper/ms.tex
update paper

% BEGIN NUMBERS
\newcommand{\numberpcs}{29\ }  % 2019/06/28 unanimous gold
\newcommand{\sVInumberlcs}{67{,}601\ }  % 2019/06/20 not allnan
\newcommand{\sVIInumberlcs}{91{,}736\ }  % 2019/06/20 not allnan
\newcommand{\numberlcs}{159{,}337\ } % 2019/06/20 sum
\newcommand{\numberclusters}{596\ } % 2019/07/01 paper_plot_all_figures.py
\newcommand{\pctoflcswithage}{77} %  2019/07/02 from hist_logt.png

\newcommand{\stscilink} {\url{archive.stsci.edu/prepds/cdips}}

@@

\begin{abstract}
  % Context
  The Transiting Exoplanet Survey Satellite (TESS) has begun to
  photometrically monitor almost every star cluster in
  the solar neighborhood.
  % Aims
  As part of the Cluster Difference Imaging Photometric Survey
  (CDIPS), we have been processing the TESS full-frame images and
  making light curves of stars that could be members of stellar
  associations, moving groups, or open clusters.
  %
  Our eventual hope is to discover giant transiting planets with known
  ages, and probe the exoplanet size and separation distributions for
  planets younger than 1 Gyr.
  %
  In this study, we made \numberlcs light curves of candidate young
  stars, across \numberclusters distinct clusters.
  Each light curve represents between 20 and 25 days of observations
  of a star brighter than $G_{Rp}=16$, sampled at 30 minute cadence.
  % Methods
  We describe in detail the image subtraction and time-series analysis 
  techniques we used to create the light curves, which
  have noise properties that agree with theoretical expectations.
  % Results
  We then highlight the \numberpcs planet candidates we identified in
  an initial search, describe the vetting procedures that they have
  survived, and emphasize the additional spectroscopic, photometric,
  and imaging observations that are needed to verify (i) whether these
  candidates are planets and (ii) whether the host stars are cluster members.
  %
  We also comment on the possible utility of the light curve sample for
  studies of stellar rotation evolution, and binary eccentricity
  damping.
  %
  The light curves and vetting reports are available at
  \stscilink.
\end{abstract}

@@

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
    %\includegraphics[width=0.45\textwidth]{target_star_cumulative_counts.png}
    \gridline{\fig{target_star_cumulative_counts.png}{0.41\textwidth}{}}
    \vspace{-0.8cm}
    \gridline{\fig{target_star_hist_logt.png}{0.41\textwidth}{}}
    \vspace{-0.8cm}
    \gridline{\fig{target_star_reference_pie_chart.png}{0.41\textwidth}{}}
	\end{center}
	\vspace{-0.8cm}
	\caption{
    Target star statistics.
    {\it Top.} Cumulative counts as a function of Gaia $Rp$-band
    magnitude.  The stars are either candidate members of clusters, or
    else have other youth indicators (see \S~\ref{sec:starselection}).
    {\it Middle.} Histogram of target star ages, for the subset of
    stars with ages matched against \citet{Kharchenko_et_al_2013}.
    {\it Bottom.} Provenance of cluster membership.  Percentages are
    relative to $N_{\rm total}=1061446$ target stars. Symbols
    are as follows.
    D14 is \citet{dias_proper_2014}.
    K13 is \citet{Kharchenko_et_al_2013}.
    CG18 is \citet{cantat-gaudin_gaia_2018}.
    Z18 is \citet{zari_3d_2018}, with upper main sequence and
    pre-main-sequence samples sub-divided.
    ``$\geq 2$'' indicates at least two authors reported a star as a
    candidate cluster member.
    %``Other'' refers to sources enumerated in \S~\ref{sec:starselection}.
	\label{fig:cdips_targets}
	}
\end{figure}

\begin{figure*}[!ht]
	\gridline{\fig{mwscmatchstats.png}{0.95\textwidth}{}}
	\vspace{-0.8cm}
	\gridline{\fig{dias14matchstats.png}{0.95\textwidth}{}}
	\vspace{-0.8cm}
	\caption{
    {\it Top.} Quality diagnostics from cross-matching
    \cite{Kharchenko_et_al_2013} cluster members against Gaia-DR2.
    histogram of the distances between matched stars is on the left; a
    histogram of the difference between the true $G$-band magnitude
    and that predicted from 2MASS photometry is in the middle; a scatter
    plot of the same magnitude difference as a function of $G$-band
    magnitude is on the right.
    {\it Bottom.} Same, but cross-matching \cite{dias_proper_2014}
    cluster members to Gaia-DR2.
	}
	\label{fig:xmatch_info}
\end{figure*}

@@

First, we use \texttt{fistar} to perform source extraction on bright
stars in each image.  During the initial source extraction, we also
fit elongated gaussians to the bright stars, yielding the shape
parameters $(s,d,k)$, where the flux $f$ as a function of position
$(x,y)$ in the CCD image plane is assumed to take the form
\begin{align}
  f(x,y) &= B + A \exp \{ -0.5 \times [
    s(\Delta x^2 + \Delta y^2) + \\
    \nonumber
    &d(\Delta x^2 - \Delta y^2) +
    k(2\Delta x \Delta y)
  ]  \},
\end{align}
for $(x_0,y_0)$ arbitrary reference positions,
$\Delta x = x-x_0$, $\Delta y = y - y_0$, and $A$ and $B$ arbitrary
flux-scaling constants.
For a nearly circular shape profile, the sharpness $s$ is related to
the FWHM as ${\rm FWHM} \approx 2.35\sqrt{s}$ \citep[{\it
e.g.},][]{Pal_2009}.  These shape parameters are later used when
selecting an astrometric reference frame
(\S~\ref{subsec:imagesubtraction}).  
In agreement with what is obvious upon visual inspection, this 
fitting process shows that stars closer to the
center of each camera's field are round, while
stars near the field edges are more triangular.

@@

\subsubsection{History of image subtraction method}
The core operation of ``classical'' image subtraction
attempts to match a photometric reference $R$ and a target image $I$ by 
computing and applying a convolution kernel.
For ground-based data, this ``match'' typically corrects for
differences in seeing or transparency between the reference and
target; for space-based data, the match might correct for spacecraft
jitter, or thermal and corresponding point-spread function (PSF)
variations.
The kernel, once applied to the high signal-to noise reference,
produces a model image, $M_{xy}$,
\begin{equation}
    M_{xy} = (R \otimes K)_{xy} + B_{xy},
    \label{eq:imagemodel}
\end{equation}
where $B_{xy}$ is a component of the model image that allows for
background variations.
Since we modeled the background separately (\S~\ref{subsec:preparation}),
we set $B_{xy}=0$.
The convolution kernel $K$ is typically decomposed onto a
basis, $K = \sum_i c_i K_i$, and the coefficients are found
by minimizing
\begin{equation}
    \chi^2 = \sum_{xy} \left( \frac{I_{xy} - M_{xy}}{\sigma_{xy}} \right)^2,
    \label{eq:chisq_conv}
\end{equation}
where $\sigma_{xy}$ is the uncertainty in the target image pixel values.
Photometry is then performed on the difference image $D_{xy}$, where
$D_{xy} = I_{xy} - M_{xy}$.
For the present reduction, the uncertainty in each target image pixel was
taken to be a constant.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{epdparams_vs_time.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		Timeseries of ``external parameters'' for a representative
			star. {\it Top}: Instrumental raw magnitude (with a particular
		aperture size), as a function of time.  {\it Second and third from
			top}: $x$ and $y$ centroid positions as a function of time.
		Continuing in order are the CCD temperature, the $(s,d,k)$ shape
		parameters, and the measured background value.
		Most of the apparent variability is instrumental.
		\label{fig:external_parameter_timeseries}
	}
\end{figure}

@@

Some notes on our implementation follow.
We select template stars in two stages.
In the first stage, we fit a parabola in the RMS-magnitude
plane, and discard stars more than $2\sigma$ away from the
prediction of the fit.
We also require that these initial candidate stars have intermediate
brightness ($8.5 > T > 13$), and have a relatively large number
of time-series data points.
We then perform an initial iteration of TFA, on only the candidate
template stars.
We inspect the resulting detrended light curves for residual
structure by computing a Lomb-Scargle periodogram.
If the maximum-power peak has a false alarm probability below 0.1\%, 
we exclude the star from the list of candidate template stars, on the
basis of its presumed periodic variability.
We then randomly select at most 200 template stars from the remaining
non-variable candidates. 
The choice of number of template stars was discussed by
\citet{kovacs_trend_2005}, and is another free parameter in
the broad problem of light curve production.
This choice is analogous to the issue of how many cotrending basis
vectors to choose in PDC-MAP or ARC2 \citep{aigrain_robust_2017}.
While the number of template stars can be optimized by constructing
and minimizing a BIC-like quantity, a little overfitting is acceptable
for our pupose of finding planetary transits.

Once the template stars are selected, we use the \texttt{VARTOOLS}
program to perform the detrending \citep{Hartman_Bakos_2016}.
This process is performed for each photometric aperture
separately.

@@

\paragraph{Sector 7}
In Sector 7, \sVIInumberlcs light curves of candidate cluster stars
were made. The provenance of the claimed cluster origin of these
sources is  ...

The clusters with the largest claimed numbers of sources are ...

@@

The youngest sample of stars is the reddest, as might be expected
given that many of these stars have not yet reached the main sequence,
and may exist along lines of sight with heavy extinction.

%FIXME
The weird part is that the oldest stars seem to have relatively more
sub-day peak periods.

@@

Maybe JH or JNW could contribute these subsections?

@@

@@

%FIXME: paraphrase / improve all the below
In this study, we made \numberlcs light curves of candidate young
stars, across \numberclusters distinct clusters.
Each light curve represents between 20 and 25 days of observations
of a star brighter than $G_{Rp}=16$, sampled at 30 minute cadence.

In the process of making the light curves, we learned X, Y, and Z.
One point in particular was that image subtraction analyses seem to
do fine at detecting variability in the galactic plane.

In analyzing the noise properties of our light curves, we found that
at the faint end, contamination from unresolved background stars
dominated over other noise sources.

From a cursory transiting planet search, we identified \numberpcs
planet candidates, which need additional spectroscopic, photometric,
and imaging observations to verify (i) whether these candidates are
planets and (ii) whether the host stars are cluster members.
The vetting reports are available at \stscilink, and the planet
candidate parameters are on exoFOP-TESS.

We also briefly explored the age-evolution of extremely high-SNR
variability in the space of dominant period versus color
(Figure~\ref{fig:ls_peak_vs_time}).
This seems to show X, and Y.

The overall takeaway is that the sample of light curves and vetting
reports we have made should prove useful for doing some astrophysics.
%  We also comment on the possible utility of the light curve sample for
%  studies of stellar rotation evolution, and binary eccentricity
%  damping.

>>>>

% BEGIN NUMBERS
\newcommand{\numberpcs}{39\ }  % 2019/06/28 unanimous gold
\newcommand{\numberzaripcs}{10\ }  % 2019/06/28 unanimous gold
\newcommand{\numberclusterpcs}{29\ }  % 2019/06/28 unanimous gold
\newcommand{\sVInumberlcs}{67{,}601\ }  % 2019/06/20 not allnan
\newcommand{\sVIInumberlcs}{91{,}736\ }  % 2019/06/20 not allnan
\newcommand{\numberlcs}{159{,}337\ } % 2019/06/20 sum
\newcommand{\numberclusters}{596\ } % 2019/07/01 paper_plot_all_figures.py
\newcommand{\pctoflcswithage}{77} %  2019/07/02 from hist_logt.png

\newcommand{\stscilink}{\url{archive.stsci.edu/prepds/cdips}}
\newcommand{\stscivetlink}{\url{archive.stsci.edu/prepds/cdips/vetting}}

@@

\begin{abstract}
  % Context
  The Transiting Exoplanet Survey Satellite (TESS) has begun to
  photometrically monitor almost every star cluster in
  the solar neighborhood.
  % Aims
  To make light curves of stars that could be members of stellar
  associations, moving groups, or open clusters, we have begun a
  Cluster Difference Imaging Photometric Survey (CDIPS).
  %
  Our aim is to discover giant transiting planets with known
  ages, and to probe the exoplanet size and separation distributions for
  planets younger than one gigayear.
  %
  In this study, we made \numberlcs light curves of candidate young
  stars, across \numberclusters distinct clusters.
  Each light curve represents between 20 and 25 days of observations
  of a star brighter than $G_{Rp}=16$, sampled at 30 minute cadence.
  % Methods
  We describe in detail the image subtraction and time-series analysis 
  techniques we used to create the light curves, which
  have noise properties that agree with theoretical expectations.
  % Results
  We then highlight the \numberpcs planet candidates we identified in
  an initial search.
  \numberclusterpcs of these planet candidates are candidate cluter
  members, and the remaining \numberzaripcs have indications of youth
  relative to field stars.
  We describe the vetting procedures that these objects have survived,
  and emphasize the additional spectroscopic, photometric, and imaging
  observations that are needed to determine (i) whether these candidates
  are planets and (ii) for which cases the host stars are cluster members.
  %
  We also comment on the possible utility of the light curve sample for
  studies of stellar rotation evolution, and binary eccentricity
  damping.
  %
  The light curves and vetting reports are available at
  \stscilink.
\end{abstract}

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		%\includegraphics[width=0.45\textwidth]{target_star_cumulative_counts.png}
		\gridline{\fig{target_star_cumulative_counts.png}{0.41\textwidth}{}}
		\vspace{-0.8cm}
		\gridline{\fig{target_star_hist_logt.png}{0.41\textwidth}{}}
		\vspace{-0.8cm}
		\gridline{\fig{target_star_reference_pie_chart.png}{0.41\textwidth}{}}
	\end{center}
	\vspace{-0.8cm}
	\caption{
		Target star statistics.
		{\it Top.} Cumulative counts as a function of Gaia $Rp$-band
		magnitude.  The stars are either candidate members of clusters, or
		else have other youth indicators (see \S~\ref{sec:starselection}).
		{\it Middle.} Histogram of target star ages, for the subset of
		stars with ages matched against \citet{Kharchenko_et_al_2013}.
		{\it Bottom.} Provenance of cluster membership.  Percentages are
		relative to $N_{\rm total}=1061446$ target stars. Symbols
		are as follows.
		D14 is \citet{dias_proper_2014}.
		K13 is \citet{Kharchenko_et_al_2013}.
		CG18 is \citet{cantat-gaudin_gaia_2018}.
		Z18 is \citet{zari_3d_2018}, with upper main sequence and
		pre-main-sequence samples sub-divided.
		``$\geq 2$'' indicates at least two authors reported a star as a
		candidate cluster member.
		%``Other'' refers to sources enumerated in \S~\ref{sec:starselection}.
		\label{fig:cdips_targets}
	}
\end{figure}

\begin{figure*}[!t]
	\gridline{\fig{mwscmatchstats.png}{0.95\textwidth}{}}
	\vspace{-0.8cm}
	\gridline{\fig{dias14matchstats.png}{0.95\textwidth}{}}
	\vspace{-0.8cm}
	\caption{
		{\it Top.} Quality diagnostics from cross-matching
		\cite{Kharchenko_et_al_2013} cluster members against Gaia-DR2.
		histogram of the distances between matched stars is on the left; a
		histogram of the difference between the true $G$-band magnitude
		and that predicted from 2MASS photometry is in the middle; a scatter
		plot of the same magnitude difference as a function of $G$-band
		magnitude is on the right.
		{\it Bottom.} Same, but cross-matching \cite{dias_proper_2014}
		cluster members to Gaia-DR2.
	}
	\label{fig:xmatch_info}
\end{figure*}

@@

@@

First, we use \texttt{fistar} to perform source extraction on bright
stars in each image.  During the initial source extraction, we also
fit elongated gaussians to the bright stars, yielding the shape
parameters $(s,d,k)$, where the flux $f$ as a function of position
$(x,y)$ in the CCD image plane is assumed to take the form
\begin{align}
  f(x,y) &= B + A \exp \{ -0.5 \times [
    s(\Delta x^2 + \Delta y^2) + \\
    \nonumber
    &d(\Delta x^2 - \Delta y^2) +
    k(2\Delta x \Delta y)
  ]  \},
\end{align}
for $(x_0,y_0)$ arbitrary reference positions,
$\Delta x = x-x_0$, $\Delta y = y - y_0$, $B$ is the background
level, and $A$ is an arbitrary
flux-scaling constant.
For a nearly circular shape profile, the sharpness $s$ is related to
the FWHM as ${\rm FWHM} \approx 2.35\sqrt{s}$ \citep[{\it
e.g.},][]{Pal_2009}.  These shape parameters are later used when
selecting an astrometric reference frame
(\S~\ref{subsec:imagesubtraction}).  
In agreement with what is obvious upon visual inspection, this 
fitting process shows that stars closer to the
center of each camera's field are round, while
stars near the field edges are more triangular.

@@

\subsubsection{History of image subtraction method}
The core operation of ``classical'' image subtraction
attempts to match a photometric reference $R$ and a target image $I$ by 
computing and applying a convolution kernel.
For ground-based data, this ``match'' typically corrects for
differences in seeing or transparency between the reference and
target; for space-based data, the match might correct for spacecraft
jitter, or thermal and corresponding point-spread function (PSF)
variations.
The kernel, once applied to the high signal-to noise reference,
produces a model image, $M_{xy}$,
\begin{equation}
    M_{xy} = (R \otimes K)_{xy} + B_{xy},
    \label{eq:imagemodel}
\end{equation}
where $B_{xy}$ is a component of the model image that allows for
background variations, and $\otimes$ denotes convolution.
Since we modeled the background separately (\S~\ref{subsec:preparation}),
we set $B_{xy}=0$.
The convolution kernel $K$ is typically decomposed onto a
basis, $K = \sum_i c_i K_i$, and the coefficients are found
by minimizing
\begin{equation}
    \chi^2 = \sum_{xy} \left( \frac{I_{xy} - M_{xy}}{\sigma_{xy}} \right)^2,
    \label{eq:chisq_conv}
\end{equation}
where $\sigma_{xy}$ is the uncertainty in the target image pixel values.
Photometry is then performed on the difference image $D_{xy}$, where
$D_{xy} = I_{xy} - M_{xy}$.
For the present reduction, the uncertainty in each target image pixel was
taken to be a constant.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{epdparams_vs_time.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		Timeseries of ``external parameters'' for a representative
			star over one sector.
			{\it Top}: Instrumental raw magnitude (with a particular
		aperture size), as a function of time.  {\it Second and third from
			top}: $x$ and $y$ centroid positions as a function of time.
		Continuing in order are the CCD temperature, the $(s,d,k)$ shape
		parameters, and the measured background value.
		Most of the apparent variability is instrumental.
		\label{fig:external_parameter_timeseries}
	}
\end{figure}

@@

Some implementation notes follow.
We selected template stars in two stages.
In the first stage, we fitted a parabola in the RMS-magnitude
plane, and discarded stars more than $2\sigma$ away from the
prediction of the fit.
We also required that these initial candidate stars have intermediate
brightness ($8.5 > T > 13$), and have a relatively large number
of time-series data points.
We then performed an initial iteration of TFA, on only the candidate
template stars.
We inspected the resulting detrended light curves for residual
structure by computing a Lomb-Scargle periodogram.
If the maximum-power peak had a false alarm probability below 0.1\%, 
we excluded the star from the list of candidate template stars, on the
basis of its presumed periodic variability.
We then randomly selected at most 200 template stars from the remaining
non-variable candidates. 
The choice of number of template stars was discussed by
\citet{kovacs_trend_2005}, and is another free parameter in
the broad problem of light curve production.
This choice is analogous to the issue of how many cotrending basis
vectors to choose in PDC-MAP or ARC2 \citep{aigrain_robust_2017}.
While the number of template stars can be optimized by constructing
and minimizing a BIC-like quantity, a little overfitting is acceptable
for our pupose of finding planetary transits.

Once the template stars were selected, we use the \texttt{VARTOOLS}
program to perform the detrending \citep{Hartman_Bakos_2016}.
This process was performed for each photometric aperture
separately.

@@

\paragraph{Sector 7}
In Sector 7, \sVIInumberlcs light curves of candidate cluster stars
were made. The provenance of the claimed cluster origin of these
sources is  ... LGBTODO

The clusters with the largest claimed numbers of sources are ...
LGBTODO

@@

The youngest sample of stars is the reddest, as might be expected
given that many of these stars have not yet reached the main sequence,
and may exist along lines of sight with heavy extinction.
Stars younger than about 200 Myr (the first two age bins) also tend
to be bluer.
The oldest stars, between about 450 Myr and 1 Gyr, show a hint in
the detected periods of being ``bottom-heavy'', {\it i.e.}, there are
relatively more short-period objects in this sample.
The reason why is unclear~--~the effect could be due to systematic
differences in the stellar samples, rather than being astrophysical.
More investigation will be required.

@@

Maybe JH or JNW could work on these subsections?

@@

JH?

@@

%FIXME: paraphrase / improve all the below
In this study, we first collected an all-sky sample of about one million
candidate young stars brighter than $G_{\rm Rp}$ of 16.
82\% of the target star sample resides in a ``cluster'', a generic
term we adopted for open clusters, stellar associations, and moving
groups.
The remaining stars have photometric or astrometric indications of
their youth, and either reside on the pre-main-sequence or upper
main sequence.

We then reduced TESS full frame images taken over the course of about
two months (Sectors 6 and 7; the first fields very close to the galactic
plane).
We performed difference imaging in order to minimize the complications
of the crowded background.
Using forced aperture photometry, we made light curves for all stars
brighter than $G_{\rm Rp}$ of 13, and went three magnitudes deeper for
our target star sample.
This yielded \numberlcs light curves of candidate young stars across
\numberclusters distinct clusters.

The light curves seem to be limited in precision by photon noise at
the bright end, and by unresolved background stars at the faint end.
Though the raw light curves show significant red noise, decorrelating
against a set of template stars led to an ensemble of light curves
with very nearly gaussian noise properties.

An initial search for transiting planets, along with subsequent
vetting to remove obvious non-cluster members and photometric blends,
has revealed \numberpcs planet candidates.
\numberclusterpcs of these candidates are candidate cluter members,
and the remaining \numberzaripcs have indications of youth, relative
to field stars.
The latter sub-sample is the less interesting one for this work, but
may be of interest for other studies.
Substantial follow-up efforts on these objects are needed to determine
(i) whether the candidates are planets and (ii) whether the host stars
are truly cluster members.
We suspect that a sizeable fraction of our planet candidates may
be eclipsing binaries, however it likely also includes real planets.
The vetting reports are available at \stscivetlink, and also documented
in Appendix~\ref{appendix:vetreport}.
The planet candidate parameters are on exoFOP-TESS.

We also briefly explored other forms of photometric variability in the
dataset.
The age-evolution of extremely high-SNR variability seems to show some
evolution in the space of LS peak-period versus color
(Figure~\ref{fig:ls_peak_vs_time}), but interpreting this result will
require a more careful exploration.
In searching for transiting planets, we found many eclipsing binaries,
as well as stellar rotation and pulsation signals.
Those who wish to explore the time-evolution in these and any other
systems are invited to explore the data at \stscilink.
